---
title: "LAScatalog"
---

```{r, echo = FALSE, warnings = FALSE}
library(rgl)

r3dDefaults <- rgl::r3dDefaults
m <- structure(c(0.921, -0.146, 0.362, 0, 0.386, 0.482, -0.787, 0, 
                -0.06, 0.864, 0.5, 0, 0, 0, 0, 1), .Dim = c(4L, 4L))
r3dDefaults$FOV <- 50
r3dDefaults$userMatrix <- m
r3dDefaults$zoom <- 0.75

knitr::opts_chunk$set(
  comment = "#>", 
  collapse = TRUE,
  fig.align = "center")

rgl::setupKnitr(autoprint = TRUE) 
options(lidR.progress = FALSE)
```

## Relevant resources:

-   [Code](https://github.com/tgoodbody/lidRtutorial/blob/master/R/07_engine.R)
-   [lidRbook section](https://r-lidar.github.io/lidRbook/engine.html)

## Overview

This code performs various operations on LiDAR data using LAScatalog functionality. We visualize and inspect the data, validate the files, clip the data based on specific coordinates, generate a Canopy Height Model (CHM), compute above ground biomass, detect treetops, specify processing options, and use parallel computing.

## Environment

```{r clear_warnings, warnings = FALSE, message=FALSE}
# Clear environment
rm(list = ls(globalenv()))

# Load packages
library(lidR)
library(sf)
```

## Basic Usage

In this section, we will cover the basic usage of the `lidR` package, including reading LiDAR data, visualization, and inspecting metadata.

### Read catalog from directory of files

We begin by creating a LAS catalog (`ctg`) from a folder containing multiple LAS files using the `readLAScatalog()` function.

```{r read_catalog}
# Read catalog and drop withheld
ctg <- readLAScatalog(folder = "data/Farm_A/",filter = "-drop_withheld")
```

### Inspect catalog

We can inspect the contents of the catalog using standard R functions.

```{r inspect_catalog}
ctg
```

### Visualize catalog

We visualize the catalog, showing the spatial coverage of the LiDAR data *header* extents. The map can be interactive if we use `map = TRUE`. Try clicking on a tile to see its header information.

```{r visualize_catalog}
plot(ctg)

# Interactive
plot(ctg, map = TRUE)
```

### Check coordinate system and extent info

We examine the coordinate system and extent information of the catalog.

```{r coordinate_system}
# coordinate system
crs(ctg)
projection(ctg)
st_crs(ctg)

# spatial extents
extent(ctg)
bbox(ctg)
st_bbox(ctg)
```

## Validate files in catalog

We validate the LAS files within the catalog using the `las_check` function. It works the same way as it would on a regular `LAS` file.

```{r validate_files}
# Perform a check on the catalog
las_check(las = ctg)
```

## File indexing

We explore indexing of `LAScatalog` objects for efficient processing.

::: callout-note
## Indexing

**The `lidR` policy has always been: use `LAStools` and `lasindex` for spatial indexing**. If you really don't want, or can't use `LAStools`, then there is a hidden function in `lidR` that users can use (`lidR:::catalog_laxindex()`).
:::

```{r, echo = FALSE, results = FALSE}
# Instructions for cleaning up any existing .lax files
# (Note: Please replace 'path' with the appropriate path)
path <- "data/Farm_A/"
file_list <- list.files(path)
delete_lax <- file_list[grep("\\.lax$", file_list)]
file.remove(file.path(path, delete_lax))
```

```{r indexing_files}
# check if files have .lax
is.indexed(ctg)

# generate index files
lidR:::catalog_laxindex(ctg)

# check if files have .lax
is.indexed(ctg)
```

## Generate CHM

We create a CHM by rasterizing the point cloud data from the catalog.

```{r generate_chm, warning = FALSE, message = FALSE}
# Generate CHM
chm <- rasterize_canopy(las = ctg, res = 0.5, algorithm = p2r(subcircle = 0.15))
plot(chm, col = height.colors(50))
```

We encounter issues and warnings while generating the CHM. Let's figure out how to fix the warnings and get decent outputs.

```{r chm_issues}
# Check for warnings
warnings()
```

## Catalog processing options

We explore and manipulate catalog options.

```{r catalog_options, warning = FALSE, message = FALSE}
# Setting options and re-rasterizing the CHM
opt_filter(ctg) <- "-drop_z_below 0 -drop_z_above 40"
opt_select(ctg) <- "xyz"
chm <- rasterize_canopy(las = ctg, res = 0.5, algorithm = p2r(subcircle = 0.15))
plot(chm, col = height.colors(50))
```

## Area-based approach on catalog

In this section, we generate Above Ground Biomass (ABA) estimates using the LAScatalog.

### Generate ABA output and visualize

We calculate ABA using the `pixel_metrics` function and visualize the results.

```{r generate_aba, warning = FALSE, message = FALSE}
# Generate area-based metrics
model <- pixel_metrics(las = ctg, func = ~max(Z), res = 20)
plot(model, col = height.colors(50))
```

## First returns only

We adjust the catalog options to calculate ABA based on first returns only.

```{r generate_aba_first_returns, warning = FALSE, message = FALSE}
opt_filter(ctg) <- "-drop_z_below 0 -drop_z_above 40 -keep_first"
model <- pixel_metrics(las = ctg, func = ~max(Z), res = 20)
plot(model, col = height.colors(50))
```

## Clip a catalog

We clip the `LAS` data in the catalog using specified coordinate groups.

```{r clip_las, warning = FALSE, message = FALSE}
# Set coordinate groups
x <- c(207846, 208131, 208010, 207852, 207400)
y <- c(7357315, 7357537, 7357372, 7357548, 7357900)

# Visualize coordinate groups
plot(ctg)
points(x, y)

# Clip plots
rois <- clip_circle(las = ctg, xcenter = x, ycenter = y, radius = 30)
```

``` r
plot(rois[[1]])
```

```{r visualize_clipped_ctg, echo = FALSE, rgl = TRUE, fig.width = 8, fig.height = 6}
# Visualize the LiDAR data with a default color palette
plot(rois[[1]], bg = "white")
```

``` r
plot(rois[[3]])
```

```{r visualize_clipped_ctg_two, echo = FALSE, rgl = TRUE, fig.width = 8, fig.height = 6}
# Visualize the LiDAR data with a default color palette
plot(rois[[3]], bg = "white")
```

## Validate clipped data

We validate the clipped LAS data using the `las_check` function.

```{r validate_clipped}
las_check(rois[[1]])
las_check(rois[[3]])
```

## Independent files (e.g. plots) as catalogs

We read an individual LAS file as a catalog and perform operations on it.

```{r delete_dems, echo = FALSE, results = FALSE}
# Instructions for cleaning up any existing .lax files
# (Note: Please replace 'path' with the appropriate path)
path <- paste0(tempdir())
file_list <- list.files(path)
delete_tif <- file_list[grep("\\.tif$", file_list)]
delete_las <- file_list[grep("\\.laz$", file_list)]
file.remove(file.path(path, delete_tif))
file.remove(file.path(path, delete_las))
```

```{r independent_files, warning = FALSE, message = FALSE}
# Read single file as catalog
ctg_non_norm <- readLAScatalog(folder = "data/MixedEucaNat.laz")

# Set options for output files
opt_output_files(ctg_non_norm) <- paste0(tempdir(),"/{XCENTER}_{XCENTER}")

# Write file as .laz
opt_laz_compression(ctg_non_norm) <- TRUE

# Get random plot locations and clip
x <- runif(n = 4, min = ctg_non_norm$Min.X, max = ctg_non_norm$Max.X)
y <- runif(n = 4, min = ctg_non_norm$Min.Y, max = ctg_non_norm$Max.Y)
rois <- clip_circle(las = ctg_non_norm, xcenter = x, ycenter = y, radius = 10)
```

```{r plots_catalog, warning = FALSE, message = FALSE}
# Read catalog of plots
ctg_plots <- readLAScatalog(tempdir())

# Set independent files option
opt_independent_files(ctg_plots) <- TRUE
opt_output_files(ctg_plots) <- paste0(tempdir(),"/{XCENTER}_{XCENTER}")

# Generate plot-level terrain models
rasterize_terrain(las = ctg_plots, res = 1, algorithm = tin())
```

```{r check_plot}
# Check files
path <- paste0(tempdir())
file_list <- list.files(path, full.names = TRUE)
file <- file_list[grep("\\.tif$", file_list)][[1]]

# plot dtm
plot(terra::rast(file))
```

## ITD using `LAScatalog`

In this section, we explore Individual Tree Detection (ITD) using the `LAScatalog`. We first configure catalog options for ITD.

```{r set_catalog_options_itd}
# Set catalog options
opt_filter(ctg) <- "-drop_withheld -drop_z_below 0 -drop_z_above 40"
```

### Detect treetops and visualize

We detect treetops and visualize the results.

```{r detect_treetops, warning = FALSE, message = FALSE}
# Detect tree tops and plot
ttops <- locate_trees(las = ctg, algorithm = lmf(ws = 3, hmin = 5))
plot(chm, col = height.colors(50))
plot(ttops, add = TRUE, cex = 0.1, col = "black")
```

### Specify catalog options

We specify additional catalog options for ITD.

```{r specify_catalog_options_itd, warning = FALSE, message = FALSE}
# Specify more options
opt_select(ctg) <- "xyz"
opt_chunk_size(ctg) <- 300
opt_chunk_buffer(ctg) <- 10

# Detect treetops and plot
ttops <- locate_trees(las = ctg, algorithm = lmf(ws = 3, hmin = 5))
plot(chm, col = height.colors(50))
plot(ttops, add = TRUE, cex = 0.1, col = "black")
```

### Parallel computing

In this section, we explore parallel computing using the `lidR` package.

## Load `future` library

We load the `future` library to enable parallel processing.

```{r load_future}
library(future)
```

## Specify catalog options

We specify catalog options for parallel processing.

```{r specify_catalog_options_parallel}
# Specify options
opt_select(ctg) <- "xyz"
opt_chunk_size(ctg) <- 300
opt_chunk_buffer(ctg) <- 10

# Visualize and summarize the catalog chunks
plot(ctg, chunk = TRUE)
summary(ctg)
```

## Single core processing

We perform tree detection using a single core.

```{r single_core_processing, warning = FALSE, message = FALSE}
# Process on single core
future::plan(sequential)

# Detect trees
ttops <- locate_trees(las = ctg, algorithm = lmf(ws = 3, hmin = 5))
```

## Parallel processing

We perform tree detection using multiple cores in parallel.

```{r parallel_processing, warning = FALSE, message = FALSE}
# Process multi-core
future::plan(multisession)

# Detect trees
ttops <- locate_trees(las = ctg, algorithm = lmf(ws = 3, hmin = 5))
```

## Revert to single core

We revert to single core processing using `future::plan(sequential)`.

```{r revert_to_single_core}
# Back to single core
future::plan(sequential)
```

This concludes the tutorial on basic usage, catalog validation, indexing, CHM generation, ABA estimation, data clipping, ITD using catalog, and parallel computing using the `lidR` package in R.

## Exercises and Questions

::: callout-tip
This exercise is complex because it involves options not yet described. Be sure to use the [lidRbook](https://r-lidar.github.io/lidRbook/) and [package documentation](https://cran.r-project.org/web/packages/lidR/lidR.pdf).
:::

Using:

`ctg <- readLAScatalog(folder = "data/Farm_A/")`

#### E1.

Generate a raster of point density for the provided catalog. Hint: Look through the documentation for a function that will do this!

#### E2.

Modify the catalog to have a point density of 10 pts/m2 using the `decimate_points()` function. If you get an error make sure to read the documentation for `decimate_points()` and try: using `opt_output_file()` to write files to a temporary directory.

#### E3.

Generate a raster of point density for this new decimated dataset.

#### E4.

Read the whole decimated catalog as a single `las` file. The catalog isn't very big - not recommended for larger datasets!

#### E5.

Read documentation for the `catalog_retile()` function and merge the decimated catalog into larger tiles.
